---
title: "Replication: The Structure of Inequality and the Politics of Redistribution"
author: "Filippo Teoldi, Zara Riaz and Julian Gerez"
date: "October 23rd, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Design declaration

We start by loading in the `DeclareDesign` package and defining the elements of the design.

* `declare_population` refers to the sample size of the study. The study concerns country-year units. In this case, there are 858 observations.
* `declare_potential_oucomes` refers to

```{r}
library('DeclareDesign')
library('panelAR')
```

## Declare Design

* Model (M): The paper tests the social affinity hypothesis that predicts middle-income voters will support redistribution policies when the distance between the middle and poor is small relative to the distance between the middle and the affluent (skew). (pg. 317)

* Inquiry (I): The study attempts to estimate parameter

$\beta$ from the simple model

$y_i = \alpha + \beta x_i + \epsilon$. 

This can be conceptualized as the summary of all potential outcomes across conditions $\beta$ for all units. More specifically, this would be all potential outcomes of redistribution conditional on inequality measured in a variety of ways, such as skew. We assume that this model describes the true data generating process.

* Data Strategy (D): The study uses country-year data on redistribution (% change in Gini coefficients brought about by taxes and government transfers) and non-elderly social spending for 18 OECD countries for the period 1969-2005. The data is purely observational and does not involve sampling. 

* Answer Strategy (A): The authors use a variety of different answer strategies. To simplify the design declaration, we focus on their basic model specification, which can be written as:

$$
R_{i,t} = \alpha + \beta X_{i,t} + \gamma R_{i, t-1} + \epsilon_{i,t}
$$

Where for each country $i$ and year $t$, $R_{i,t}$ refers to redistribution, $\alpha$ is the constant, $\beta$ is the coefficient for a measure of inequality (e.g. we are using skew and the 9010 ratiobelow), $\gamma$ is the coefficient for redistribution lagged by one year ($R_{i,t-1}$) and  $\epsilon_{i,t}$ is the error term.

```{r}
# We model the primary specifications used to measure redistribution outcomes (Table 2, Column 7)
#Step 1: Load actual data
load("redistsample.Rdata")
data<- redistsample

#Step 2: Generate parameters
modelX <- panelAR(redist ~ dvpratio9050 + dvpratio5010 + as.factor(id),
                data=redistsample, panelVar='id', timeVar='time', autoCorr='ar1',
                panelCorrMethod='pcse',rho.na.rm=TRUE, panel.weight='t-1',
                bound.rho=TRUE)
summary(modelX)
library(lmtest)
results<- coeftest(modelX) #storing coefficients from model
results
a_X <- results["(Intercept)","Estimate"]
b_X <- results["dvpratio9050","Estimate"]
c_X <- results["dvpratio5010", "Estimate"]

# Create vector of residuals
u_X <- rep(NA, 87)
names(u_X) <- rownames(redistsample)
u_X[names(modelX$residuals)] <- modelX$residuals

country_coef <- results[grepl("id", names(results[,1])), "Estimate"] # get all country fixed effects

# Create vectors of the length of the entire dataset so that each country has correct FE
country_fe <- rep(c(0, country_coef), times = as.numeric(table(redistsample$id)))
```

```{r}
# Step 1: M -- Model: Speculation on variables and relations between them
population <- declare_population(
  country = add_level(
    N = 18,  country_fe = rnorm(N)),
 year = add_level(
    N = 45, t = 1:N, nest = F),
  obs = cross_levels(
    by = join(country, year),
    dvpratio9050 = rnorm(n = N, mean=1.0388, sd=0.176),
    dvpratio5010 = rnorm(n = N, mean=1.731, sd=0.285),
    u_i = country_fe + rnorm(N)/2
  )
)
pop_frame<- population()

#reduce <- function(data) dplyr::filter(())
```

```{r}

modelX <- panelAR(redist ~ dvpratio9050 + dvpratio5010 + as.factor(id),
                data=redistsample, panelVar='id', timeVar='time', autoCorr='ar1',
                panelCorrMethod='pcse',rho.na.rm=TRUE, panel.weight='t-1',
                bound.rho=TRUE)
summary(modelX)
library(lmtest)
results<- coeftest(modelX) #storing coefficients from model
results
a_X <- results["(Intercept)","Estimate"]
b_X <- results["dvpratio9050","Estimate"]
c_X <- results["dvpratio5010", "Estimate"]

# Create vector of residuals
u_X <- rep(NA, 87)
names(u_X) <- rownames(redistsample)
u_X[names(modelX$residuals)] <- modelX$residuals

country_coef <- results[grepl("id", names(results[,1])), "Estimate"] # get all country fixed effects

# Step 2: M -- Model: Speculation on variables and relations between them
fx <- function(X, Z, e) {
  redist <- b_X*X + c_X*Z + e 
  return(redist)
}

lagged_dv <- function(redistsample, lambda = .9) {
  within(redistsample, {Y <- NA
                Y[t == 1] <- 1 + u_i[t == 1]
                for(j in 2:max(t))  {Y[t==j] <- (1-lambda) + lambda*Y[t==(j - 1)] + u_i[t==j]}
                })
                 }

add_dv <- declare_step(handler = lagged_dv)


outcomes<- declare_step(handler=fabricate,
          redist1= fx(dvpratio9050, dvpratio5010, u_i)           
          )
```


```{r}
# Step 4 - Create Missingness - first go is easy and MAR
miss_function <- function(redistsample, m=771){ # m is number of missing cases
  redistsample[sort(sample(1:nrow(redistsample), nrow(redistsample)-m)),]
}

missingness <- declare_step(miss_function)


```

```{r}
  estimand <- declare_estimand(
  SPO= mean((fx(max(dvpratio9050/dvpratio5010), country_fe, u_X)) -
                   fx(min(dvpratio9050/dvpratio5010),country_fe, u_X))/
                   (max(dvpratio9050/dvpratio5010)-min(dvpratio9050/dvpratio5010)))
```
# D -- Data Strategy: No random sampling given the study uses observational data

```{r}
estimator_1 <- declare_estimator(redist1 ~ (dvpratio9050/dvpratio5010), country_fe
                                 , estimand = "SPO", 
                                  model = panelAR,
                                 timeVar="t",
                                 panelVar= "country",
                                 autoCorr='ar1',
                                  panelCorrMethod='pcse',
                                 rhotype="breg", 
                                 rho.na.rm=TRUE,
                                 panel.weight='t-1',
                          
                                 label = "panelAR")


```

```{r}
lupu_pontusson_2011_design <- population  + outcomes + missingness + estimand + estimator_1

temp_design <- population + outcomes + estimand

temp_design <- population  + outcomes + missingness + estimand
library(dplyr)
df <- population()  %>% outcomes %>% missingness %>% estimand
head(population())
outcomes()

```
---
#Diagnosis
```{r}
library(knitr)

plot(redistsample$redist, redistsample$dvskew)
plot(pop_frame$redist1, pop_frame$dvpratio9050)

kable(reshape_diagnosis(diagnose_design(temp_design, sims = 500)))

diagnosis <- diagnose_design(
  design = lupu_pontusson_2011_design, 
  sims = 500
)

kable(reshape_diagnosis((diagnosis)))

```