---
title: "Extension and Robustness Checks: The Structure of Inequality and the Politics of Redistribution"
author: "Filippo Teoldi, Zara Riaz and Julian Gerez"
date: "October 23rd, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Extension

```{r}
library('haven')
# Loading data that contains disaggregated data for disaggregated categories of social spending
disag_data <- read_dta("disag_spending.dta")

# Imputing missing values using `mice`
library('zoo')

disagdata_countries <- lapply(unique(disag_data$country), function(x)
  subset(disag_data, disag_data$country==x)
)

library(mice)
data_imputed<-complete(mice(disag_data,m=5,maxit=50,meth='pmm',seed=500))
```

## Merging and cleaning merged data
```{r}
load("data_socspend.Rdata")
merged <- merge(data_socspend, data_imputed, by=c("country","year"))


#Creating lags for each dependent variable
merged$fambenefits_lag <- unlist(by(merged ,merged$id, 
                                    function(x){c(NA,x[,"fambenefits"][1:(length(x[,"fambenefits"])-1)])}))

merged$incapacity_lag <- unlist(by(merged ,merged$id, 
                                   function(x){c(NA,x[,"incapacity"][1:(length(x[,"incapacity"])-1)])}))

merged$pubspendinglbr_lag <- unlist(by(merged ,merged$id, 
                                       function(x){c(NA,x[,"pubspending_labor"][1:(length(x[,"pubspending_labor"])-1)])}))

merged$pubunemplag <- unlist(by(merged ,merged$id,
                                function(x){c(NA,x[,"public_unemp"][1:(length(x[,"public_unemp"])-1)])}))

#Setting up time-series conditions on merged data:

merged <- merged[with(merged , order(id, year)),]

merged$time <- unlist(by(merged ,merged$id,function(x) seq(1:nrow(x))))
```

## Running models from table 3 on disaggregated outcomes

```{r}
library('panelAR')
#Main model (using 90/50 and 50/10 ratios)
ext1a <- panelAR(fambenefits ~ fambenefits_lag + ma_pratio9050s + ma_pratio5010s + ma_pop65 +
                  ma_turnout + ma_fempar + ma_stddisp_gall + ma_pvoc + ma_union +
                  ma_unempl + ma_dreher + gdpgrowth, 
                data=merged , panelVar='id', timeVar='time', 
                autoCorr='psar1', panelCorrMethod='pcse',
                rho.na.rm=TRUE, bound.rho=TRUE)
summary(ext1a)


ext1b <- panelAR(incapacity ~ incapacity_lag + ma_pratio9050s + ma_pratio5010s + ma_pop65 +
                  ma_turnout + ma_fempar + ma_stddisp_gall + ma_pvoc + ma_union +
                  ma_unempl + ma_dreher + gdpgrowth, 
                data=merged , panelVar='id', timeVar='time', 
                autoCorr='psar1', panelCorrMethod='pcse',
                rho.na.rm=TRUE, bound.rho=TRUE)
summary(ext1b)

ext1c <- panelAR(pubspending_labor ~ pubspendinglbr_lag + ma_pratio9050s + ma_pratio5010s + ma_pop65 +
                  ma_turnout + ma_fempar + ma_stddisp_gall + ma_pvoc + ma_union +
                  ma_unempl + ma_dreher + gdpgrowth, 
                data=merged , panelVar='id', timeVar='time', 
                autoCorr='psar1', panelCorrMethod='pcse',
                rho.na.rm=TRUE, bound.rho=TRUE)
summary(ext1c)

ext1d <- panelAR(public_unemp ~ pubunemplag + ma_pratio9050s + ma_pratio5010s + ma_pop65 +
                  ma_turnout + ma_fempar + ma_stddisp_gall + ma_pvoc + ma_union +
                  ma_unempl + ma_dreher + gdpgrowth, 
                data=merged , panelVar='id', timeVar='time', 
                autoCorr='psar1', panelCorrMethod='pcse',
                rho.na.rm=TRUE, bound.rho=TRUE)
summary(ext1d)

##Model without controls

ext2a <- panelAR(fambenefits ~ ma_pratio9050s + ma_pratio5010s +gdpgrowth + as.factor(id), data=merged , panelVar='id', timeVar='time', autoCorr='psar1', panelCorrMethod='pcse',rho.na.rm=TRUE, panel.weight='t-1', bound.rho=TRUE)
summary(ext2a)

ext2b <- panelAR(incapacity ~ ma_pratio9050s + ma_pratio5010s +gdpgrowth + as.factor(id), data=merged , panelVar='id', timeVar='time', autoCorr='psar1', panelCorrMethod='pcse',rho.na.rm=TRUE, panel.weight='t-1', bound.rho=TRUE)
summary(ext2b)

ext2c <- panelAR(pubspending_labor ~ ma_pratio9050s + ma_pratio5010s +gdpgrowth + as.factor(id), data=merged , panelVar='id', timeVar='time', autoCorr='psar1', panelCorrMethod='pcse',rho.na.rm=TRUE, panel.weight='t-1', bound.rho=TRUE)
summary(ext2c)

ext2d <- panelAR(public_unemp ~ ma_pratio9050s + ma_pratio5010s +gdpgrowth + as.factor(id), data=merged , panelVar='id', timeVar='time', autoCorr='psar1', panelCorrMethod='pcse',rho.na.rm=TRUE, panel.weight='t-1', bound.rho=TRUE)
summary(ext2d)

#Using skew and 90/10 ratios:

ext3a <- panelAR(fambenefits ~ fambenefits_lag + ma_skews + ma_pratio9010s + ma_pop65 +
                  ma_turnout + ma_fempar + ma_stddisp_gall + ma_pvoc + ma_union +
                  ma_unempl + ma_dreher + gdpgrowth, 
                data=merged , panelVar='id', timeVar='time', 
                autoCorr='psar1', panelCorrMethod='pcse',
                rho.na.rm=TRUE, bound.rho=TRUE)
summary(ext3a)


ext3b <- panelAR(incapacity ~ incapacity_lag + ma_skews + ma_pratio9010s + ma_pop65 +
                  ma_turnout + ma_fempar + ma_stddisp_gall + ma_pvoc + ma_union +
                  ma_unempl + ma_dreher + gdpgrowth, 
                data=merged , panelVar='id', timeVar='time', 
                autoCorr='psar1', panelCorrMethod='pcse',
                rho.na.rm=TRUE, bound.rho=TRUE)
summary(ext3b)

ext3c <- panelAR(pubspending_labor ~ pubspendinglbr_lag + ma_skews + ma_pratio9010s + ma_pop65 +
                  ma_turnout + ma_fempar + ma_stddisp_gall + ma_pvoc + ma_union +
                  ma_unempl + ma_dreher + gdpgrowth, 
                data=merged , panelVar='id', timeVar='time', 
                autoCorr='psar1', panelCorrMethod='pcse',
                rho.na.rm=TRUE, bound.rho=TRUE)
summary(ext3c)

ext3d <- panelAR(public_unemp ~ pubunemplag + ma_skews + ma_pratio9010s + ma_pop65 +
                  ma_turnout + ma_fempar + ma_stddisp_gall + ma_pvoc + ma_union +
                  ma_unempl + ma_dreher + gdpgrowth, 
                data=merged , panelVar='id', timeVar='time', 
                autoCorr='psar1', panelCorrMethod='pcse',
                rho.na.rm=TRUE, bound.rho=TRUE)
summary(ext3d)

#Using skew and ratio and no other controls

ext4a <- panelAR(fambenefits ~  ma_skews + ma_pratio9010s + gdpgrowth, 
                data=merged , panelVar='id', timeVar='time', 
                autoCorr='psar1', panelCorrMethod='pcse',
                rho.na.rm=TRUE, bound.rho=TRUE)
summary(ext4a)


ext4b <- panelAR(incapacity ~ ma_skews + ma_pratio9010s + gdpgrowth,  
                data=merged , panelVar='id', timeVar='time', 
                autoCorr='psar1', panelCorrMethod='pcse',
                rho.na.rm=TRUE, bound.rho=TRUE)
summary(ext4b)

ext4c <- panelAR(pubspending_labor ~ ma_skews + ma_pratio9010s + gdpgrowth, 
                data=merged , panelVar='id', timeVar='time', 
                autoCorr='psar1', panelCorrMethod='pcse',
                rho.na.rm=TRUE, bound.rho=TRUE)
summary(ext4c)

ext4d <- panelAR(public_unemp ~ ma_skews + ma_pratio9010s + gdpgrowth, 
                data=merged , panelVar='id', timeVar='time', 
                autoCorr='psar1', panelCorrMethod='pcse',
                rho.na.rm=TRUE, bound.rho=TRUE)
summary(ext4d)
```

# Robustness checks

```{r}
load("redistsample.Rdata")
```

## Table 2a: AR(1)-type autocorrelation with a unique correlation coefficient for each panel

### Specification 1

```{r}
mod1_out1 <- panelAR(redist ~ redist_lag + dvpratio9050 + dvpratio5010 + dvturnout +
                  dvfempar + dvstddisp_gall + dvpvoc + dvunion + dvunempl,
                data=redistsample, panelVar='id', timeVar='time', autoCorr='psar1',
                panelCorrMethod='pcse',rho.na.rm=TRUE, panel.weight='t-1',
                bound.rho=TRUE)
summary(mod1_out1)
```

### Specification 2 (remove outliers):

```{r}
# defining outliers 
mod1.resid <- mod1_out1$residuals
index <- which(abs((mod1.resid-mean(mod1.resid))/sd(mod1.resid)) <= 1.5)
# creating a new subset without these observations
redistsample_noout<- mod1_out1$model[index,]

# running same model as spec1 with new subset
mod1_out2 <- panelAR(redist ~ redist_lag + dvpratio9050 + dvpratio5010 + dvturnout +
                  dvfempar + dvstddisp_gall + dvpvoc + dvunion + dvunempl,
                data=redistsample_noout, panelVar='id', timeVar='time', 
                autoCorr='psar1', panelCorrMethod='pcse',rho.na.rm=TRUE,
                panel.weight='t-1', bound.rho=TRUE)
summary(mod1_out2)
```

### Specification 3 (no controls, no country dummies):

```{r}
mod1_out3 <- panelAR(redist ~ dvpratio9050 + dvpratio5010,
                data=redistsample, panelVar='id', timeVar='time', autoCorr='psar1',
                panelCorrMethod='pcse',rho.na.rm=TRUE, panel.weight='t-1',
                bound.rho=TRUE)
summary(mod1_out3)
```

### Specification 4 (no controls, no outliers, no country dummies):

```{r}
# defining outliers 
mod3.resid <- mod1_out3$residuals
index <- which(abs((mod3.resid-mean(mod3.resid))/sd(mod3.resid)) <= 1.5)
# creating a new subset without these observations
redistsample_noout<- mod1_out3$model[index,]
# running same model as spec3 with new subset
mod1_out4 <- panelAR(redist ~ dvpratio9050 + dvpratio5010,
                data=redistsample_noout, panelVar='id', timeVar='time',
                autoCorr='psar1', panelCorrMethod='pcse',rho.na.rm=TRUE,
                panel.weight='t-1', bound.rho=TRUE)
summary(mod1_out4)
```

### Specification 5 (using skew as main inequality measure):

```{r}
mod1_out5 <- panelAR(redist ~ redist_lag + dvratio9010 + dvskew + dvturnout + dvfempar +
                 dvstddisp_gall + dvpvoc + dvunion + dvunempl, data=redistsample,
               panelVar='id', timeVar='time', autoCorr='psar1',
               panelCorrMethod='pcse',rho.na.rm=TRUE, panel.weight='t-1',
               bound.rho=TRUE) 
summary(mod1_out5)
```

### Specification 6 (skew as main measure, no outliers):

```{r}
mod5.resid <- mod1_out5$residuals
index <- which(abs((mod5.resid-mean(mod5.resid))/sd(mod5.resid)) <= 1.5)
#creating a new subset without these observations
redistsample_noout<- mod1_out5$model[index,]
#running same model as spec5 with new subset
mod1_out6 <- panelAR(redist ~ redist_lag + dvratio9010 + dvskew + dvturnout + dvfempar +
                 dvstddisp_gall + dvpvoc + dvunion + dvunempl, 
               data=redistsample_noout, panelVar='id', timeVar='time', 
               autoCorr='psar1', panelCorrMethod='pcse',rho.na.rm=TRUE,
               panel.weight='t-1', bound.rho=TRUE) 
summary(mod1_out6)
```

### Specification 7 (skew as main measure, no controls, no country dummies):

```{r}
mod1_out7 <- panelAR(redist ~ dvratio9010 + dvskew, 
                data=redistsample, panelVar='id', timeVar='time',
                autoCorr='psar1', panelCorrMethod='pcse',rho.na.rm=TRUE,
                panel.weight='t-1', bound.rho=TRUE)
summary(mod1_out7)
```

### Specification 8 (skew as main measure, no controls, no country dummies without outliers):

```{r}
mod7.resid <- mod1_out7$residuals
index <- which(abs((mod7.resid-mean(mod7.resid))/sd(mod7.resid)) <= 1.5)
#creating a new subset without these observations
redistsample_noout<- mod1_out7$model[index,]
#running same model as spec7 with new subset
mod1_out8 <- panelAR(redist ~ dvratio9010 + dvskew, 
                data=redistsample_noout, panelVar='id', timeVar='time',
                autoCorr='psar1', panelCorrMethod='pcse',rho.na.rm=TRUE,
                panel.weight='t-1', bound.rho=TRUE)
summary(mod1_out8)
```

## Table 2b: different panel correction methods (a = Huber-White sandiwch estimator, b = panel-weighted least squares)

### Specification 1

```{r}
mod2a_out1 <- panelAR(redist ~ redist_lag + dvpratio9050 + dvpratio5010 + dvturnout +
                  dvfempar + dvstddisp_gall + dvpvoc + dvunion + dvunempl,
                data=redistsample, panelVar='id', timeVar='time', autoCorr='ar1',
                panelCorrMethod='phet',rho.na.rm=TRUE, panel.weight='t-1',
                bound.rho=TRUE)
summary(mod2a_out1)

mod2b_out1 <- panelAR(redist ~ redist_lag + dvpratio9050 + dvpratio5010 + dvturnout +
                  dvfempar + dvstddisp_gall + dvpvoc + dvunion + dvunempl,
                data=redistsample, panelVar='id', timeVar='time', autoCorr='ar1',
                panelCorrMethod='pwls',rho.na.rm=TRUE, panel.weight='t-1',
                bound.rho=TRUE)
summary(mod2b_out1)
```

### Specification 2 (remove outliers):

```{r}
# defining outliers 
mod1.resid <- mod2a_out1$residuals
index <- which(abs((mod1.resid-mean(mod1.resid))/sd(mod1.resid)) <= 1.5)
# creating a new subset without these observations
redistsample_noout<- mod2a_out1$model[index,]

# running same model as spec1 with new subset
mod2a_out2 <- panelAR(redist ~ redist_lag + dvpratio9050 + dvpratio5010 + dvturnout +
                  dvfempar + dvstddisp_gall + dvpvoc + dvunion + dvunempl,
                data=redistsample_noout, panelVar='id', timeVar='time', 
                autoCorr='ar1', panelCorrMethod='phet',rho.na.rm=TRUE,
                panel.weight='t-1', bound.rho=TRUE)
summary(mod2a_out2)

# defining outliers 
mod1.resid <- mod2b_out1$residuals
index <- which(abs((mod1.resid-mean(mod1.resid))/sd(mod1.resid)) <= 1.5)
# creating a new subset without these observations
redistsample_noout<- mod2b_out1$model[index,]

# running same model as spec1 with new subset
mod2b_out2 <- panelAR(redist ~ redist_lag + dvpratio9050 + dvpratio5010 + dvturnout +
                  dvfempar + dvstddisp_gall + dvpvoc + dvunion + dvunempl,
                data=redistsample_noout, panelVar='id', timeVar='time', 
                autoCorr='ar1', panelCorrMethod='pwls',rho.na.rm=TRUE,
                panel.weight='t-1', bound.rho=TRUE)
summary(mod2b_out2)
```

### Specification 3 (no controls):

```{r}
mod2a_out3 <- panelAR(redist ~ dvpratio9050 + dvpratio5010 + as.factor(id),
                data=redistsample, panelVar='id', timeVar='time', autoCorr='ar1',
                panelCorrMethod='phet',rho.na.rm=TRUE, panel.weight='t-1',
                bound.rho=TRUE)
summary(mod2a_out3)

# no country dummies
mod2b_out3 <- panelAR(redist ~ dvpratio9050 + dvpratio5010,
                data=redistsample, panelVar='id', timeVar='time', autoCorr='ar1',
                panelCorrMethod='pwls',rho.na.rm=TRUE, panel.weight='t-1',
                bound.rho=TRUE)
summary(mod2b_out3)
```

### Specification 4 (no controls, no outliers):

```{r}
# defining outliers 
mod3.resid <- mod2a_out3$residuals
index <- which(abs((mod3.resid-mean(mod3.resid))/sd(mod3.resid)) <= 1.5)
# creating a new subset without these observations
redistsample_noout<- mod2a_out3$model[index,]
# running same model as spec3 with new subset
mod2a_out4 <- panelAR(redist ~ dvpratio9050 + dvpratio5010 + as.factor(id),
                data=redistsample_noout, panelVar='id', timeVar='time',
                autoCorr='ar1', panelCorrMethod='phet',rho.na.rm=TRUE,
                panel.weight='t-1', bound.rho=TRUE)
summary(mod2a_out4)

# no country fixed effects

# defining outliers 
mod3.resid <- mod2b_out3$residuals
index <- which(abs((mod3.resid-mean(mod3.resid))/sd(mod3.resid)) <= 1.5)
# creating a new subset without these observations
redistsample_noout<- mod2a_out3$model[index,]
# running same model as spec3 with new subset
mod2b_out4 <- panelAR(redist ~ dvpratio9050 + dvpratio5010,
                data=redistsample_noout, panelVar='id', timeVar='time',
                autoCorr='ar1', panelCorrMethod='pwls',rho.na.rm=TRUE,
                panel.weight='t-1', bound.rho=TRUE)
summary(mod2b_out4)
```

### Specification 5 (using skew as main inequality measure):

```{r}
mod2a_out5<- panelAR(redist ~ redist_lag + dvratio9010 + dvskew + dvturnout + dvfempar +
                 dvstddisp_gall + dvpvoc + dvunion + dvunempl, data=redistsample,
               panelVar='id', timeVar='time', autoCorr='ar1',
               panelCorrMethod='phet',rho.na.rm=TRUE, panel.weight='t-1',
               bound.rho=TRUE) 
summary(mod2a_out5)

mod2b_out5<- panelAR(redist ~ redist_lag + dvratio9010 + dvskew + dvturnout + dvfempar +
                 dvstddisp_gall + dvpvoc + dvunion + dvunempl, data=redistsample,
               panelVar='id', timeVar='time', autoCorr='ar1',
               panelCorrMethod='pwls',rho.na.rm=TRUE, panel.weight='t-1',
               bound.rho=TRUE) 
summary(mod2b_out5)
```

### Specification 6 (skew as main measure, no outliers):

```{r}
mod5.resid <- mod2a_out5$residuals
index <- which(abs((mod5.resid-mean(mod5.resid))/sd(mod5.resid)) <= 1.5)
#creating a new subset without these observations
redistsample_noout<- mod2a_out5$model[index,]
#running same model as spec5 with new subset
mod2a_out6<- panelAR(redist ~ redist_lag + dvratio9010 + dvskew + dvturnout + dvfempar +
                 dvstddisp_gall + dvpvoc + dvunion + dvunempl, 
               data=redistsample_noout, panelVar='id', timeVar='time', 
               autoCorr='ar1', panelCorrMethod='phet',rho.na.rm=TRUE,
               panel.weight='t-1', bound.rho=TRUE) 
summary(mod2a_out6)

mod5.resid <- mod2b_out5$residuals
index <- which(abs((mod5.resid-mean(mod5.resid))/sd(mod5.resid)) <= 1.5)
#creating a new subset without these observations
redistsample_noout<- mod2b_out5$model[index,]
#running same model as spec5 with new subset
mod2b_out6<- panelAR(redist ~ redist_lag + dvratio9010 + dvskew + dvturnout + dvfempar +
                 dvstddisp_gall + dvpvoc + dvunion + dvunempl, 
               data=redistsample_noout, panelVar='id', timeVar='time', 
               autoCorr='ar1', panelCorrMethod='pwls',rho.na.rm=TRUE,
               panel.weight='t-1', bound.rho=TRUE) 
summary(mod2b_out6)
```

### Specification 7 (skew as main measure, no controls, country dummies):

```{r}
mod2a_out7 <- panelAR(redist ~ dvratio9010 + dvskew + as.factor(id), 
                data=redistsample, panelVar='id', timeVar='time',
                autoCorr='ar1', panelCorrMethod='phet',rho.na.rm=TRUE,
                panel.weight='t-1', bound.rho=TRUE)
summary(mod2a_out7)

# no country dummies
mod2b_out7 <- panelAR(redist ~ dvratio9010 + dvskew, 
                data=redistsample, panelVar='id', timeVar='time',
                autoCorr='ar1', panelCorrMethod='pwls',rho.na.rm=TRUE,
                panel.weight='t-1', bound.rho=TRUE)
summary(mod2b_out7)
```

### Specification 8 (skew as main measure, no controls, fixed effects without outliers):

```{r}
mod7.resid <- mod2a_out7$residuals
index <- which(abs((mod7.resid-mean(mod7.resid))/sd(mod7.resid)) <= 1.5)
#creating a new subset without these observations
redistsample_noout<- mod2a_out7$model[index,]
#running same model as spec7 with new subset
mod2a_out8 <- panelAR(redist ~ dvratio9010 + dvskew + as.factor(id), 
                data=redistsample_noout, panelVar='id', timeVar='time',
                autoCorr='ar1', panelCorrMethod='phet',rho.na.rm=TRUE,
                panel.weight='t-1', bound.rho=TRUE)
summary(mod2a_out8)

# no country dummies

mod7.resid <- mod2a_out7$residuals
index <- which(abs((mod7.resid-mean(mod7.resid))/sd(mod7.resid)) <= 1.5)
#creating a new subset without these observations
redistsample_noout<- mod2a_out7$model[index,]
#running same model as spec7 with new subset
mod2a_out8 <- panelAR(redist ~ dvratio9010 + dvskew, 
                data=redistsample_noout, panelVar='id', timeVar='time',
                autoCorr='ar1', panelCorrMethod='pwls',rho.na.rm=TRUE,
                panel.weight='t-1', bound.rho=TRUE)
summary(mod2a_out8)
```