---
title: "Extension and Robustness Checks: The Structure of Inequality and the Politics of Redistribution"
author: "Filippo Teoldi, Zara Riaz and Julian Gerez"
date: "October 23rd, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
set.seed(2172)
```

# Extension: Disaggregating Social Spending Outcomes

```{r}
library('haven')
# Loading data that contains disaggregated data for disaggregated categories of social spending
disag_data <- read_dta("disag_spending.dta")

# Imputing missing values using `mice`
library('zoo')

disagdata_countries <- lapply(unique(disag_data$country), function(x)
  subset(disag_data, disag_data$country==x)
)

library('mice')
data_imputed<-complete(mice(disag_data,m=5,maxit=50,meth='pmm',seed=500))
```

## Merging and cleaning merged data
```{r}
load("data_socspend.Rdata")
merged <- merge(data_socspend, data_imputed, by=c("country","year"))

#Creating lags for each dependent variable
merged$fambenefits_lag <- unlist(by(merged ,merged$id, 
                                    function(x){c(NA,x[,"fambenefits"][1:(length(x[,"fambenefits"])-1)])}))

merged$incapacity_lag <- unlist(by(merged ,merged$id, 
                                   function(x){c(NA,x[,"incapacity"][1:(length(x[,"incapacity"])-1)])}))

merged$pubspendinglbr_lag <- unlist(by(merged ,merged$id, 
                                       function(x){c(NA,x[,"pubspending_labor"][1:(length(x[,"pubspending_labor"])-1)])}))

merged$pubunemplag <- unlist(by(merged ,merged$id,
                                function(x){c(NA,x[,"public_unemp"][1:(length(x[,"public_unemp"])-1)])}))

#Setting up time-series conditions on merged data:

merged <- merged[with(merged , order(id, year)),]

merged$time <- unlist(by(merged ,merged$id,function(x) seq(1:nrow(x))))
```

## Running models from table 3 on disaggregated outcomes

```{r}
library('panelAR')
#Main model (using 90/50 and 50/10 ratios)
ext1a <- panelAR(fambenefits ~ fambenefits_lag + ma_pratio9050s + ma_pratio5010s + ma_pop65 +
                  ma_turnout + ma_fempar + ma_stddisp_gall + ma_pvoc + ma_union +
                  ma_unempl + ma_dreher + gdpgrowth, 
                data=merged , panelVar='id', timeVar='time', 
                autoCorr='psar1', panelCorrMethod='pcse',
                rho.na.rm=TRUE, bound.rho=TRUE)
summary(ext1a)


ext1b <- panelAR(incapacity ~ incapacity_lag + ma_pratio9050s + ma_pratio5010s + ma_pop65 +
                  ma_turnout + ma_fempar + ma_stddisp_gall + ma_pvoc + ma_union +
                  ma_unempl + ma_dreher + gdpgrowth, 
                data=merged , panelVar='id', timeVar='time', 
                autoCorr='psar1', panelCorrMethod='pcse',
                rho.na.rm=TRUE, bound.rho=TRUE)
summary(ext1b)

ext1c <- panelAR(pubspending_labor ~ pubspendinglbr_lag + ma_pratio9050s + ma_pratio5010s + ma_pop65 +
                  ma_turnout + ma_fempar + ma_stddisp_gall + ma_pvoc + ma_union +
                  ma_unempl + ma_dreher + gdpgrowth, 
                data=merged , panelVar='id', timeVar='time', 
                autoCorr='psar1', panelCorrMethod='pcse',
                rho.na.rm=TRUE, bound.rho=TRUE)
summary(ext1c)

ext1d <- panelAR(public_unemp ~ pubunemplag + ma_pratio9050s + ma_pratio5010s + ma_pop65 +
                  ma_turnout + ma_fempar + ma_stddisp_gall + ma_pvoc + ma_union +
                  ma_unempl + ma_dreher + gdpgrowth, 
                data=merged , panelVar='id', timeVar='time', 
                autoCorr='psar1', panelCorrMethod='pcse',
                rho.na.rm=TRUE, bound.rho=TRUE)
summary(ext1d)

#Using skew and 90/10 ratios:

ext2a <- panelAR(fambenefits ~ fambenefits_lag + ma_skews + ma_pratio9010s + ma_pop65 +
                  ma_turnout + ma_fempar + ma_stddisp_gall + ma_pvoc + ma_union +
                  ma_unempl + ma_dreher + gdpgrowth, 
                data=merged , panelVar='id', timeVar='time', 
                autoCorr='psar1', panelCorrMethod='pcse',
                rho.na.rm=TRUE, bound.rho=TRUE)
summary(ext2a)


ext2b <- panelAR(incapacity ~ incapacity_lag + ma_skews + ma_pratio9010s + ma_pop65 +
                  ma_turnout + ma_fempar + ma_stddisp_gall + ma_pvoc + ma_union +
                  ma_unempl + ma_dreher + gdpgrowth, 
                data=merged , panelVar='id', timeVar='time', 
                autoCorr='psar1', panelCorrMethod='pcse',
                rho.na.rm=TRUE, bound.rho=TRUE)
summary(ext2b)

ext2c <- panelAR(pubspending_labor ~ pubspendinglbr_lag + ma_skews + ma_pratio9010s + ma_pop65 +
                  ma_turnout + ma_fempar + ma_stddisp_gall + ma_pvoc + ma_union +
                  ma_unempl + ma_dreher + gdpgrowth, 
                data=merged , panelVar='id', timeVar='time', 
                autoCorr='psar1', panelCorrMethod='pcse',
                rho.na.rm=TRUE, bound.rho=TRUE)
summary(ext2c)

ext2d <- panelAR(public_unemp ~ pubunemplag + ma_skews + ma_pratio9010s + ma_pop65 +
                  ma_turnout + ma_fempar + ma_stddisp_gall + ma_pvoc + ma_union +
                  ma_unempl + ma_dreher + gdpgrowth, 
                data=merged , panelVar='id', timeVar='time', 
                autoCorr='psar1', panelCorrMethod='pcse',
                rho.na.rm=TRUE, bound.rho=TRUE)
summary(ext2d)

```

# Robustness checks

```{r}
load("redistsample.Rdata")
```

## Table 2a: AR(1)-type autocorrelation with a unique correlation coefficient for each panel

### Specification 1

```{r}
mod1_out1 <- panelAR(redist ~ redist_lag + dvpratio9050 + dvpratio5010 + dvturnout +
                  dvfempar + dvstddisp_gall + dvpvoc + dvunion + dvunempl,
                data=redistsample, panelVar='id', timeVar='time', autoCorr='psar1',
                panelCorrMethod='pcse',rho.na.rm=TRUE, panel.weight='t-1',
                bound.rho=TRUE)
summary(mod1_out1)
```

### Specification 5 (using skew as main inequality measure):

```{r}
mod1_out5 <- panelAR(redist ~ redist_lag + dvratio9010 + dvskew + dvturnout + dvfempar +
                 dvstddisp_gall + dvpvoc + dvunion + dvunempl, data=redistsample,
               panelVar='id', timeVar='time', autoCorr='psar1',
               panelCorrMethod='pcse',rho.na.rm=TRUE, panel.weight='t-1',
               bound.rho=TRUE) 
summary(mod1_out5)
```

## Table 2b: different panel correction methods (a = Huber-White sandiwch estimator, b = panel-weighted least squares)

### Specification 1

```{r}
mod2a_out1 <- panelAR(redist ~ redist_lag + dvpratio9050 + dvpratio5010 + dvturnout +
                  dvfempar + dvstddisp_gall + dvpvoc + dvunion + dvunempl,
                data=redistsample, panelVar='id', timeVar='time', autoCorr='ar1',
                panelCorrMethod='phet',rho.na.rm=TRUE, panel.weight='t-1',
                bound.rho=TRUE)
summary(mod2a_out1)

mod2b_out1 <- panelAR(redist ~ redist_lag + dvpratio9050 + dvpratio5010 + dvturnout +
                  dvfempar + dvstddisp_gall + dvpvoc + dvunion + dvunempl,
                data=redistsample, panelVar='id', timeVar='time', autoCorr='ar1',
                panelCorrMethod='pwls',rho.na.rm=TRUE, panel.weight='t-1',
                bound.rho=TRUE)
summary(mod2b_out1)
```

### Specification 5 (using skew as main inequality measure):

```{r}
mod2a_out5<- panelAR(redist ~ redist_lag + dvratio9010 + dvskew + dvturnout + dvfempar +
                 dvstddisp_gall + dvpvoc + dvunion + dvunempl, data=redistsample,
               panelVar='id', timeVar='time', autoCorr='ar1',
               panelCorrMethod='phet',rho.na.rm=TRUE, panel.weight='t-1',
               bound.rho=TRUE) 
summary(mod2a_out5)

mod2b_out5<- panelAR(redist ~ redist_lag + dvratio9010 + dvskew + dvturnout + dvfempar +
                 dvstddisp_gall + dvpvoc + dvunion + dvunempl, data=redistsample,
               panelVar='id', timeVar='time', autoCorr='ar1',
               panelCorrMethod='pwls',rho.na.rm=TRUE, panel.weight='t-1',
               bound.rho=TRUE) 
summary(mod2b_out5)
```


#Model Extensions/Modifications
```{r}

#OLS
load("redistsample.Rdata")
data<- redistsample

#Main model (using 90/50 and 50/10 ratios)

#install.packages("estimatr")
library(estimatr)
ols1 <- lm_robust(redist ~ redist_lag + dvpratio9050 + dvpratio5010 + dvturnout +
                  dvfempar + dvstddisp_gall + dvpvoc + dvunion + dvunempl,
                data=redistsample)
summary(ols1)


#Main model without country FEs or controls

ols2 <- lm(redist ~ dvpratio9050 + dvpratio5010 ,
                data=redistsample)
summary(ols2)
library(lmtest)
coeftest(ols2)


#Main model with country FEs and no controls

ols3 <- lm(redist ~ dvpratio9050 + dvpratio5010 + as.factor(id),
                data=redistsample)
summary(ols3)
coeftest(ols3)

#install.packages("plm")
library("plm")
library("sandwich")
waldtest(ols2, ols3, vcov = vcovHC)

# Main model replacing ratios with skew

ols4<- lm_robust(redist ~ redist_lag + dvratio9010 + dvskew + dvturnout + dvfempar +
                 dvstddisp_gall + dvpvoc + dvunion + dvunempl, data=redistsample)
summary(ols4)
coeftest(ols4)

#Main Social Spending Model

data_socspend <- data[,c(1:3, 18, 20, 48:62 )]

data_socspend  <- data_socspend [with(data_socspend , order(id, year)),]

data_socspend$socspend_lag <- unlist(by(data_socspend ,data_socspend$id,                                    function(x){c(NA,x[,"socspend"][1:(length(x[,"socspend"])-1)])}))

#Main model (using 90/50 and 50/10 ratios)
ols5<- lm_robust(socspend ~ socspend_lag + ma_pratio9050s + ma_pratio5010s + ma_pop65 +
                  ma_turnout + ma_fempar + ma_stddisp_gall + ma_pvoc + ma_union +
                  ma_unempl + ma_dreher + gdpgrowth, 
                data=data_socspend)
summary(ols5)
coeftest(ols5)

#Main model with country FEs and no controls
ols7 <- lm_robust(socspend ~ ma_pratio9050s + ma_pratio5010s +gdpgrowth, data=data_socspend)
summary(ols7)

#Main model with country FEs and no controls
ols8 <- lm_robust(socspend ~ ma_pratio9050s + ma_pratio5010s +gdpgrowth + as.factor(id), data=data_socspend)
summary(ols8)
# Main model replacing ratios with skew
ols9<- lm_robust(socspend ~ socspend_lag + ma_pratio9010s + ma_skews + ma_pop65+ ma_turnout + ma_fempar + ma_stddisp_gall + ma_pvoc + ma_union + ma_unempl + ma_dreher + gdpgrowth, data=data_socspend)
summary(ols9)

```

#Social Spending Robustness: 2 year moving-averages 
```{r}
#How slow moving is the process? -Using 2 year moving averages 

data_2yr <- cbind(data, sapply(c(20, 26:27, 32, 34, 21, 15, 22:23, 10:13, 29:30), function(b)
  (lag(data[,b], 1)+lag(data[,b], 2))/2
))

names(data_2yr)[65:79] <- c("ma2_socspend", "ma2_pratio9050s", "ma2_pratio5010s",
                        "ma2_pratio9010s", "ma2_skews", "ma2_dreher", "ma2_pop65",
                        "ma2_stdpjoint", "ma2_stddisp_gall", "ma2_fempar", "ma2_unempl",
                        "ma2_union", "ma2_turnout", "ma2_pvoc", "ma2_fpop")

data2yr_socspend <- data_2yr[,c(1:3, 18, 20, 65:79 )]

data2yr_socspend  <- data2yr_socspend [with(data_socspend , order(id, year)),]

data2yr_socspend$socspend_lag <- unlist(by(data2yr_socspend ,data2yr_socspend$id,                                    function(x){c(NA,x[,"socspend"][1:(length(x[,"socspend"])-1)])}))

data2yr_socspend$time <- unlist(by(
  data2yr_socspend ,data2yr_socspend $id,function(x) seq(1:nrow(x))))
```

#Running models with 2 year averages
```{r}

```

#Defining skew as the difference rather than ratio
```{r}
data_socspend$skew2<- (data_socspend$ma_pratio9050s)-(data_socspend$ma_pratio5010s)
redistsample$skew2<-(redistsample$dvpratio9050)-(redistsample$dvpratio5010)

#Redistribution

redistskew<- panelAR(redist ~ redist_lag + dvratio9010 + skew2 + dvturnout + dvfempar +
                 dvstddisp_gall + dvpvoc + dvunion + dvunempl, data=redistsample,
               panelVar='id', timeVar='time', autoCorr='ar1',
               panelCorrMethod='pcse',rho.na.rm=TRUE, panel.weight='t-1',
               bound.rho=TRUE) 
summary(redistskew)

#Social Spending

data_socspend$time <- unlist(by(
  data_socspend ,data_socspend $id,function(x) seq(1:nrow(x))))

socspend_skew<- panelAR(socspend ~ socspend_lag + ma_pratio9010s + skew2 + ma_pop65+ ma_turnout + ma_fempar + ma_stddisp_gall + ma_pvoc + ma_union + ma_unempl + ma_dreher + gdpgrowth, data=data_socspend, panelVar='id', timeVar='time', autoCorr='psar1', panelCorrMethod='pcse',rho.na.rm=TRUE, panel.weight='t-1', bound.rho=TRUE) 
summary(socspend_skew)

```

```{r}

twoyear1 <- panelAR(socspend ~ socspend_lag + ma2_pratio9050s + ma2_pratio5010s + ma2_pop65 +
                  ma2_turnout + ma2_fempar + ma2_stddisp_gall + ma2_pvoc + ma2_union +
                  ma2_unempl + ma2_dreher + gdpgrowth, 
                data=data2yr_socspend , panelVar='id', timeVar='time', 
                autoCorr='psar1', panelCorrMethod='pcse',
              rho.na.rm=TRUE, bound.rho=TRUE)
summary(twoyear1)

#With skew
twoyear3 <- panelAR(socspend ~ socspend_lag + ma2_pratio9010s + ma2_skews + ma2_pop65+ ma2_turnout + ma2_fempar + ma2_stddisp_gall + ma2_pvoc + ma2_union + ma2_unempl + ma2_dreher + gdpgrowth, data=data2yr_socspend, panelVar='id', timeVar='time', autoCorr='psar1', panelCorrMethod='pcse',rho.na.rm=TRUE, panel.weight='t-1', bound.rho=TRUE) 
summary(twoyear3)

```
